Large-scale harmonic patterns clustering

### 2010-03-11 Paper/Paper.tex

% RESULTS
% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 
% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 0.
% 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 
% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 
% 200codes, 3 beats, pSize=3, maxener, True, False, not scaled = 
\begin{table}
\begin{center}
\begin{tabular}{|l|c|}
\hline
size & avg. dist \\ \hline \hline
1 bar (4) & \\
2 bars (8) & $$\\
2 bars (12) & $$\\
2 bars (16) & $$ \\ \hline
3 beats & \\
5 beats & $$\\
7 beats & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Encoding score for different pattern sizes. We can select the
size of the patterns based on bars. In this case, we put the actual size
of the pattern in parentheses. If the pattern size is based on beats, the
size is the number of beats. Results are on the uspop dataset, not seen
during training.}}
\label{tab:psize}
\end{table}

\subsection{Data Size}
We are interesting to see how the amount of training data changes
the quality of our code.
We experiment with different data sizes for a fixed size of codebook and a 
fixed code length. The data points are chose randomly from the whole
data set.
Code length: 1 bar resized to size 4. Codebook size: 100
We measure the distortion on our test set. We hope for some convergence.
See results in Figure \ref{fig:sizes}.

@@

\iffalse
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
size & $5$K & $10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
distortion & $$ & $0.143$ & $0.142$ & $0.142$ & $0.142$ & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a codebook of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
\fi
% RESULTS:    
% 1bar s4     avg_distOLD    avg_dist
% 1K          0.150370
% 5K          0.145327
% 10K         0.143567
% 50K         0.142257
% 100K        0.142195
% 250K        0.141796
% 500K        0.142081
% 1M          0.142118
% 2M          0.141596
% 1bar s4 first X samples
% 1K          0.157142
% 5K          0.150045
% 10K         0.148357
% 50K         0.146069
% 100K        0.145765
% 250K        0.145819
% 500K        0.145772
% 1M          0.145924
% 2M          0.146065
% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
% 1K          0.155474
% 5K          0.149619
% 10K         0.148056
% 50K         0.146328
% 100K        0.146004
% 250K        0.146024
% 500K        0.145962
% 1M          0.145873
% 2M          0.145955
% 2bar s8     
% 1K          0.171300
% 5K          0.165960
% 10K         0.165063
% 50K         0.162439
% 100K        0.161588
% 250K        0.161910
% 500K        0.161233
% 1M          0.161057
% 2M          0.161068

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|l|c|}
\hline
\# codes & \# samples & avg. dist \\ \hline \hline
1 & $500$ & $0.066079$ \\
10 & $5$K & $0.045690$ \\
50 & $25$K & $0.038456$ \\
100 & $50$K & $0.036725$ \\
200 & $100$K & $0.033923$ \\
500 & $250$K & $$ \\
1000 & $500$K & \\ \hline
\end{tabular}
\end{center}
\caption{{Codebook size.}}
\label{tab:cbsize}
\end{table}
% RESULTS:      avg_distsOLD    avg_dist
% code 1        0.195939        0.066079
% code 10       0.161812        0.045690
% code 50       0.147472        0.038456
% code 100      0.142194        0.036725
% code 200      0.137548        0.033923
% code 500      0.131211        0.

\subsection{Codebook Size in Function of Pattern Length}
If the signal was totally random, and we were using 2 code to encode
every beat, we would need.... 4 codes for 2 beats? and 16 codes for
4 beats? and 256 codes for 8 beats?

Think about it in binary: two code to encode (0) and (1), but for
(0,0),(0,1),(1,0),(1,1) we need four codes for the same accuracy.

Reasonable size: 2 - 4 - 16 - 256. If we improve distance, music
is not random!

Then... 3 beats, 5 beats, 7 beats don't encode as well!

>>>>

% RESULTS
% 200codes, 2 bars, pSize=16, maxener, True, False, not scaled = 
% 200codes, 2 bars, pSize=8, maxener, True, False, not scaled  = 0.042432
% 200codes, 1 bars, pSize=4, maxener, True, False, not scaled  = 
% 200codes, 7 beats, pSize=7, maxener, True, False, not scaled = 
% 200codes, 5 beats, pSize=5, maxener, True, False, not scaled = 
% 200codes, 3 beats, pSize=3, maxener, True, False, not scaled = 
\begin{table}
\begin{center}
\begin{tabular}{|l|c|}
\hline
size & avg. dist \\ \hline \hline
1 bar (4) & \\
2 bars (8) & $$\\
2 bars (12) & $$\\
2 bars (16) & $$ \\ \hline
3 beats & \\
5 beats & $$\\
7 beats & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Encoding score for different pattern sizes. We can select the
size of the patterns based on bars. In this case, we put the actual size
of the pattern in parentheses. If the pattern size is based on beats, the
size is the number of beats. Results are on the uspop dataset, not seen
during training.}}
\label{tab:psize}
\end{table}

\subsection{Data Size}
We are interesting to see how the amount of training data changes
the quality of our codebook.
We experiment with different data sizes for a fixed size of codebook and a 
fixed code length. The data points are chose randomly from the whole
data set.
In the experiment a pattern size correspond to $1$ bar resized to $4$
beats, thus a pattern size of $12$x$4$. The codebook size is $100$.
We measure the distortion on our test set. We hope for some convergence.
See results in Figure \ref{fig:sizes}.

@@

\iffalse
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
size & $5$K & $10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
distortion & $$ & $0.143$ & $0.142$ & $0.142$ & $0.142$ & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a codebook of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
\fi
% RESULTS:    
% 1bar s4     avg_distOLD    avg_dist
% 1K          0.150370       0.041941
% 5K          0.145327       0.038449
% 10K         0.143567       0.036892
% 50K         0.142257       0.036722
% 100K        0.142195       
% 250K        0.141796
% 500K        0.142081
% 1M          0.142118
% 2M          0.141596
% 1bar s4 first X samples
% 1K          0.157142       
% 5K          0.150045
% 10K         0.148357
% 50K         0.146069
% 100K        0.145765
% 250K        0.145819
% 500K        0.145772
% 1M          0.145924
% 2M          0.146065       
% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
% 1K          0.155474       0.042596
% 5K          0.149619
% 10K         0.148056
% 50K         0.146328
% 100K        0.146004
% 250K        0.146024
% 500K        0.145962
% 1M          0.145873
% 2M          0.145955
% 2bar s8     
% 1K          0.171300       0.049282
% 5K          0.165960       0.046441
% 10K         0.165063       0.045507
% 50K         0.162439
% 100K        0.161588
% 250K        0.161910
% 500K        0.161233
% 1M          0.161057
% 2M          0.161068

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|l|c|}
\hline
\# codes & \# samples & avg. dist \\ \hline \hline
1 & $500$ & $0.066079$ \\
10 & $5$K & $0.045690$ \\
50 & $25$K & $0.038456$ \\
100 & $50$K & $0.036725$ \\
200 & $100$K & $0.033923$ \\
500 & $250$K & $$ \\
1000 & $500$K & \\ \hline
\end{tabular}
\end{center}
\caption{{Experiment with the codebook size. We train on patterns
corresponding to $1$ bar resized to $4$ beats. The number of training
samples is set to $500$ times the number of codes, which is sufficient
according to Figure \ref{fig:sizes}.}}
\label{tab:cbsize}
\end{table}
% RESULTS:      avg_distsOLD    avg_dist
% code 1        0.195939        0.066079
% code 10       0.161812        0.045690
% code 50       0.147472        0.038456
% code 100      0.142194        0.036725
% code 200      0.137548        0.033923
% code 500      0.131211        0.

\subsection{Codebook Size in Function of Pattern Length}
We experiment on the randomness of the \textit{Shingles}, assuming
that there is some correlation between beats and bars.
The intuition is as follow: if the signal was totally random, every
time the signal size double, the codebook size must be squared.
To illustrate this, imagine a binary array of size $1$. We need two
codes to encode perfectly: $(0)$ and $(1)$. If the binary is now of
size $2$, we need four codes: $(0,0)$, $(0,1)$, $(1,1)$ and $(1,0)$.

In Table \ref{tab:size_pattern}, we test that hypothesis. As expected,
the distortion improves with larger pattern sizes. The details
are difficult to analyze. The error jump from two beats to one bar
can probably be explained by the obvious case of sustained notes.
We also believe that smaller codebooks are more affected by their
initialization. Fortunately, we can conclude that music is not random,
which is always a good sanity check!

Then... 3 beats, 5 beats, 7 beats don't encode as well?
